TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-03-13 paper/draft.tex
improve wording

Low-mass stars (with masses $M_{*} < 0.7\,$\msun and atmospheric temperatures $T_{\rm eff} < 4000\,$K) are by far the most ubiquitous type of star, comprising $\sim 70 \%$ of the galaxy's population by number \citep{Bochanski:2010}. With nuclear fusion timescales of $\tau>10^{12}$ yr \citep{Laughlin:1997}, chemical compositions of M dwarf populations tracing the nucleosynthetic processes and interstellar mixing of heavy elements from many generations of shorter-lived, high mass stars, are a unique probe for piecing together galactic structure and evolution \citep{Bochanski:2010}.

Additionally, the low masses of M dwarfs make for easier detection of planets by radial velocity offset detections \citep{Trifonov:2018}, high planet-to-star radii ratios make for easier detection of exoplanet transits in light curve observations (e.g. \kepler, \tess), and shorter orbital periods allow for new planet discovery in less observation time than their higher mass counterparts. For these reasons, M dwarfs are the primary candidates for exoplanet detection searches, making detailed and precise knowledge of M dwarf chemical compositions key to constraining the atmospheric conditions of potentially habitable exoplanets through exoplanet spectra obtained through transit spectroscopy.

Advances in instrumentation and the lauch of several large scale spectroscopic stellar surveys launched in the past decade have dramatically increased the sample of known M dwarfs, including the \citealt{West:2011} catalog of 70,841 sources from the Sloan Digital Sky Survey (SDSS DR7; \citealt{Abazajian:2009}), and the \citealt{Yi:2014} and \citealt{Guo:2015} catalogs of 93,619 from the Large Sky Area Multi-Object Fiber Spectroscopic Telescope (LAMOST DR1; \citealt{Zhao:2012}), which each have determined spectral subtype classifications, H$\alpha$ equivalent width, and a number of molecular band index measurements. 
The launch of the (\apogee; \citealt{Majewski:2015}) survey as part of the Sloan Digital Sky Survey III mission \citealt{Eisenstein:2011} has introduced the largest sample of M dwarfs observed with high resolution spectroscopy.

Precise measurements of fundamental stellar parameters (including temperatures, masses, radii, abundances and kinematics) are key to studying M dwarfs on both galactic and planetary scales. Methods for determining these parameters... temperatures (models, photometry), masses (dynamical)

Why is there a problem determining M dwarf metallicities (and especially for high res \apogee\ spectra)? Emphasize issues: (1) models dont accurately reproduce data; (2) empirical standards are relative not absolute, and they are sparesely sampled. 
Methods for determining metallicity of M dwarfs: (1) from atmospheric models of M dwarfs, (2) empirical methods

>>>>

Low-mass stars (with masses $M_{*} < 0.7\,$\msun and atmospheric temperatures $T_{\rm eff} < 4000\,$K) are by far the most ubiquitous type of star, comprising $\sim 70 \%$ of the galaxy's population by number \citep{Bochanski:2010}. With nuclear fusion timescales of $\tau>10^{11}$ yr \citep{Laughlin:1997}, chemical compositions of M dwarf populations tracing the nucleosynthetic processes and interstellar mixing of heavy elements from many generations of shorter-lived, high mass stars, are a unique probe for piecing together galactic structure and evolution \citep{Bochanski:2010}.

Additionally, the low masses of M dwarfs make for easier detection of planets by radial velocity offset detections \citep{Trifonov:2018}, high planet-to-star radii ratios make for easier detection of exoplanet transits in light curve observations (e.g. \kepler, \tess), and shorter orbital periods allow for new planet discovery in less observation time than their higher mass counterparts. For these reasons, M dwarfs are the primary candidates for exoplanet detection searches, making detailed and precise knowledge of M dwarf chemical compositions key to constraining the atmospheric conditions of potentially habitable exoplanets.

Advances in instrumentation with the lauch of several spectroscopic surveys in the past decade have dramatically increased the sample of known M dwarfs, and our potential to study M dwarf abundances on both exoplanet and galactic scales at a large statistical level.
The largest of such catalogs include the \citealt{West:2011} sample of 70,841 sources from the Sloan Digital Sky Survey (SDSS DR7; \citealt{Abazajian:2009}), and the \citealt{Yi:2014} and \citealt{Guo:2015} catalogs of 93,619 from the Large Sky Area Multi-Object Fiber Spectroscopic Telescope (LAMOST DR1; \citealt{Zhao:2012}), which each have determined spectral subtype classifications, H$\alpha$ equivalent width, and a number of optical molecular band index measurements. The launch of the (\apogee; \citealt{Majewski:2015}) survey as part of the Sloan Digital Sky Survey III mission \citep{Eisenstein:2011} has introduced the largest sample of M dwarfs observed with high resolution spectroscopy.

Why is there a problem determining M dwarf metallicities (and especially for high res \apogee\ spectra)? Emphasize issues: (1) models dont accurately reproduce data; (2) empirical standards are relative not absolute, and they are sparesely sampled. 
Methods for determining metallicity of M dwarfs: (1) from atmospheric models of M dwarfs, (2) empirical methods: spectroscopic or photometric calibrations with FGK pairs.
